---
title: "Interpret GWAS results"
author: "Jinliang Yang"
date: "11-16-2020"
output: NULL
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
knitr::opts_knit$set(root.dir=normalizePath('../../')) 
#library(tidyverse)
```

# GWAS results from Mike:
/work/jyanglab/mmeier/HAVELOCK_BG/largedata/GWAS/out_traits_150_lowN_201109-100251
/work/jyanglab/mmeier/HAVELOCK_BG/largedata/GWAS/out_traits_150_stdN_201109-103909

```{bash}
cp /work/jyanglab/mmeier/HAV_19/largedata/GEMMA/sign_snps_annotated_stdN.csv largedata/gwas/
```


### In the table:
  - p = -log10(p_wald)
  - h2 = heritability score under stdN
  - asv_count = total ASV count in all stdN samples
  - genus = like "Genus" column but Family name is inserted where Genus is unknown


--------------------------------
# Basic STAT

```{r, eval=FALSE}
library("data.table")

gwas <- fread("largedata/gwas/sign_snps_annotated_stdN.csv")
gwas$snpid <- paste(gwas$chr, gwas$ps, sep="_")
names(gwas)[4] <- "pos"

g10 <- subset(gwas, p > 10)
```

### Quick Manhattan plot

```{r}
library("huskeR")

png("graphs/Manhatton_g10_stdN.png", width=11, height=4.25, units="in", res=300)
plot_mht(inputdf=g10, col2plot="p", jmph=FALSE, cl_file="data/chr_length_B73v4.csv", 
         ylab="-log10(p-value)",  cex = 0.6)
dev.off()
```





### At Plant genomics level

How many SNPs hit multiple ASVs?
1. 2.6 million SNP with at least one hit
2. frq > 1: 875503

```{r}
snp <- table(gwas$snpid)
snp <- data.frame(snp)
snp <- snp[order(snp$Freq, decreasing = TRUE),]

write.table(snp, "cache/snp_hit_frq_stdN.csv", sep=",", row.names=FALSE, quote=FALSE)
nrow(subset(snp, Freq > 1))
nrow(subset(snp, Freq == 20))
```

### Use a window approach

```{r}
gwas$pos <- as.numeric(as.character(gwas$pos))


winsize = 1000

gwas$winid <- paste(gwas$chr, round(gwas$pos/winsize,0), sep="_")
  
win <- data.frame(table(gwas$winid))
win <- win[order(win$Freq, decreasing = TRUE),]

```

```{r}
library("huskeR")

win$chr <- gsub("_.*", "", win$Var1)
win$pos <- gsub(".*_", "", win$Var1)
win$pos <- as.numeric(win$pos)*winsize

png("graphs/Manhatton_win1k_stdN.png", width=11, height=4.25, units="in", res=300)
plot_mht(inputdf=win, col2plot="Freq", jmph=FALSE, cl_file="data/chr_length_B73v4.csv", pl=TRUE,
         ylab="Window Frequency",  cex = 0.6)
dev.off()
```



### At Microbiome level

```{r}
m <- data.frame(table(gwas$ASV))
m <- m[order(m$Freq, decreasing = TRUE),]

subm <- subset(gwas, winid %in% "6_141845")

length(unique(subm$ASV))
length(unique(subm$Family))
length(unique(subm$Phylum))

tb <- data.frame(table(subm$ASV))
tb <- tb[order(tb$Freq, decreasing=TRUE), ]
```

Plot GWAS signal for asv_000464

```{r}
png("graphs/Manhatton_asv_asv_000228_stdN.png", width=11, height=4.25, units="in", res=300)
plot_mht(inputdf=subset(gwas, ASV %in% "asv_000464"), col2plot="p", jmph=FALSE, cl_file="data/chr_length_B73v4.csv", pl=FALSE,
         ylab="-log10(p-value)",  cex = 0.2)
dev.off()
```

Plot GWAS signal for asv_004687
```{r}
png("graphs/Manhatton_asv_004687_stdN.png", width=11, height=4.25, units="in", res=300)
plot_mht(inputdf=subset(gwas, ASV %in% "asv_004687"), col2plot="p", jmph=FALSE, cl_file="data/chr_length_B73v4.csv", pl=FALSE,
         ylab="-log10(p-value)",  cex = 0.2, hdf = data.frame(chr=2, pos=238142057))
dev.off()

png("graphs/Manhatton_asv_000323_stdN.png", width=11, height=4.25, units="in", res=300)
plot_mht(inputdf=subset(gwas, ASV %in% "asv_000323"), col2plot="p", jmph=FALSE, cl_file="data/chr_length_B73v4.csv", pl=FALSE,
         ylab="-log10(p-value)",  cex = 0.2, hdf = data.frame(chr=2, pos=238142057))
dev.off()

out <- subset(gwas, chr == 2 & pos > (238142057 - 1000000) & pos < (238146688 + 1000000))
# associated with 692 unique ASV
length(unique(out$ASV))
out1 <- subset(out, ASV %in% c("asv_000302", "asv_000323"))
```
-------------------

# 1000 Manhatton Plots

```{r}
asvs <- unique(gwas$ASV)
for(i in 1:1000){
  
  outid <- paste0("largedata/mht_stdN/", asvs[i], ".png")
  png(outid, width=11, height=4.25, units="in", res=300)
  plot_mht(inputdf=subset(gwas, ASV %in% asvs[i]), col2plot="p", 
           jmph=FALSE, cl_file="data/chr_length_B73v4.csv", pl=FALSE,
           main=asvs[i], ylab="-log10(p-value)",  cex = 0.2)
  dev.off()
}

```



# Burkholderia_ Manhatton Plots


```{r}
png("graphs/Manhatton_asv_asv_000228_stdN.png", width=11, height=4.25, units="in", res=300)
plot_mht(inputdf=subset(gwas, ASV %in% "asv_000464"), col2plot="p", jmph=FALSE, cl_file="data/chr_length_B73v4.csv", pl=FALSE,
         ylab="-log10(p-value)",  cex = 0.2)
dev.off()
```


```{r}
#library("tidyverse")

genera <- sort(unique(gwas$genus))


for(gn in genera){
  asvs <- unique(filter(gwas, genus == gn)$ASV)
  for(i in 1:length(asvs)){
    outid <- paste0("largedata/gwas/mht_stdN/", gn, "_", asvs[i], ".png")
    png(outid, width=11, height=4.25, units="in", res=300)
    plot_mht(inputdf=subset(gwas, ASV %in% asvs[i]), col2plot="p", 
             jmph=FALSE, cl_file="data/chr_length_B73v4.csv", pl=FALSE,
             main=paste(gn, asvs[i]), ylab="-log10(p-value)",  cex = 0.2)
    dev.off()
  }
  
}




```






